---
title: "LinearRegressionModel"
author: "Jared Babcock"
date: "11/11/2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("LoadingData.R")
source("DataTransformation.R")
```


We want to find the largest contributing factors to customer churn score (churn score is our response)

Get rid of city, zip code, lat long, latitude, and longitude for now since there are too many factors

Some cities are important to predicting churn score (based on city p-value) but not all of them

There are 20 rows where total charges is na, since that is a small subset of the data (0.284%) we drop those,
it is not worth imputing and possibly adding of data


some sources:
https://stackoverflow.com/questions/7980622/subset-of-rows-containing-na-missing-values-in-a-chosen-column-of-a-data-frame
https://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name

Create full model

```{r}
model.full <- lm(`Churn Score`~.,data=dat.reduced)
summary(model.full)
```

```{r}
cooks.distances <- cooks.distance(model.full)
n <- dim(dat.reduced)[1]
threshold <- 4/n

plot(cooks.distances,type='h',ylab="Cook's Distances")
abline(h=threshold,col='red')
```

We can see from this plot that we have several points with Cook's Distance above the 4/n threshold.

```{r}
above.indices <- cooks.distances[cooks.distances>threshold]

num.outliers <- length(above.indices)

num.outliers/n*100
```

2.79% of the data are outliers, so we will remove them.

```{r}
dat.nooutliers <- dat.reduced[cooks.distances<=threshold,]
```

```{r}
model.full <- lm(`Churn Score`~.,data=dat.nooutliers)
model.intercept.only <- lm(`Churn Score`~1,data=dat.nooutliers)

anova(model.full,model.intercept.only)
```

We can see from this ANOVA that the full model is statistically significantly better than a model with no predictors and only an intercept. So, we can do variable selection next.

```{r}
summary(model.full)
levels(dat.nooutliers$`Multiple Lines`)
```
We can see NAs present in our model summary, which indicates multicollinearity. First we will make sure that categorical variables are coded correctly. We will combine 'No' and 'No service' for multiple variables since it appears they are collinear.

```{r}
# Deal with Multiple Lines
dat.nooutliers$`Multiple Lines`[dat.nooutliers$`Multiple Lines` == 'No phone service'] <- 'No'
dat.nooutliers$`Multiple Lines` <- as.factor(dat.nooutliers$`Multiple Lines`)
dat.nooutliers$`Multiple Lines` <- droplevels(dat.nooutliers$`Multiple Lines`)

# Deal with Online Security
dat.nooutliers$`Online Security`[dat.nooutliers$`Online Security` == 'No internet service'] <- 'No'
dat.nooutliers$`Online Security` <- as.factor(dat.nooutliers$`Online Security`)
dat.nooutliers$`Online Security` <- droplevels(dat.nooutliers$`Online Security`)

# Deal with Online Backup
dat.nooutliers$`Online Backup`[dat.nooutliers$`Online Backup` == 'No internet service'] <- 'No'
dat.nooutliers$`Online Backup` <- as.factor(dat.nooutliers$`Online Backup`)
dat.nooutliers$`Online Backup` <- droplevels(dat.nooutliers$`Online Backup`)

# Deal with Device Protection
dat.nooutliers$`Device Protection`[dat.nooutliers$`Device Protection` == 'No internet service'] <- 'No'
dat.nooutliers$`Device Protection` <- as.factor(dat.nooutliers$`Device Protection`)
dat.nooutliers$`Device Protection` <- droplevels(dat.nooutliers$`Device Protection`)

# Deal with Tech Support
dat.nooutliers$`Tech Support`[dat.nooutliers$`Tech Support` == 'No internet service'] <- 'No'
dat.nooutliers$`Tech Support` <- as.factor(dat.nooutliers$`Tech Support`)
dat.nooutliers$`Tech Support` <- droplevels(dat.nooutliers$`Tech Support`)

# Deal with Streaming TV
dat.nooutliers$`Streaming TV`[dat.nooutliers$`Streaming TV` == 'No internet service'] <- 'No'
dat.nooutliers$`Streaming TV` <- as.factor(dat.nooutliers$`Streaming TV`)
dat.nooutliers$`Streaming TV` <- droplevels(dat.nooutliers$`Streaming TV`)

# Deal with Streaming Movies
dat.nooutliers$`Streaming Movies`[dat.nooutliers$`Streaming Movies` == 'No internet service'] <- 'No'
dat.nooutliers$`Streaming Movies` <- as.factor(dat.nooutliers$`Streaming Movies`)
dat.nooutliers$`Streaming Movies` <- droplevels(dat.nooutliers$`Streaming Movies`)
```

```{r}
model.full <- lm(`Churn Score`~.,data=dat.nooutliers)
summary(model.full)
```

Now evaluate VIF to see if we have to remove any other collinear variables.

```{r}

vif.threshold <- 1 / (1 - summary(model.full)$r.squared)

cutoff.val <- max(10,vif.threshold)
cutoff.val
```

The variables that have adjusted GVIF > 10 (GVIF^(1/(2*df))) exhibit collinearity and should be removed. Looking at the output below, we want to remove the variable Monthly Charges since it has a high adjusted VIF.
TODO: do we want GVIF or adjusted GVIF
```{r}
vif(model.full)
```

```{r}
# Monthly charges is in column 18, so remove that from the data

dat.nooutliers.reduced <- dat.nooutliers[,-18]
```

Use forward stepwise regression for model selection

```{r}
full <- lm(`Churn Score`~.,data=dat.nooutliers.reduced)
minimum <- lm(`Churn Score`~1,data=dat.nooutliers.reduced)
step(minimum, scope = list(lower=minimum, upper = full), direction = "forward")
```

Create reduced model

```{r}
model.reduced <- lm(formula = `Churn Score` ~ Contract + `Internet Service` + 
    `Total Charges` + Dependents + `Streaming Movies` + `Multiple Lines` + 
    `Streaming TV` + `Payment Method` + `Senior Citizen` + `Paperless Billing` + 
    `Online Security` + Partner, data = dat.nooutliers.reduced)
summary(model.reduced)
View(dat.nooutliers.reduced)
```

From the reduced model, we see the following predictors are significant at <0.001 significance value:
Contract, Internet Service, Total Charges, Dependents, Streaming Movies and Multiple Lines.