---
title: "Logistic regression"
author: "Marta Bras"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("LoadingData.R")
source("DataTransformation.R")
```


# Full model

## Fitting the model
```{r}
full.model <- glm(`Churn Value` ~., family = "binomial", data = dat.reduced_2)
summary(full.model)
```

## Coefficients

```{r}
coefficients <-  coef(full.model)

kable(coefficients, digits =6,  caption = "Logistic regression model - Coefficients") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```


## Model significance overall

```{r}
pc1 <- 8143.4-5628.0
pc2 <-7031-7008

#X2 test
1-pchisq(pc1,pc2)
```

To see if model is significant overall we do $null deviance - residual deviance$ and we test for significance of this difference with a $X^2$ test for $df_{nulldeviance} - df_{residual_deviance}$. Since the result from the test is very close to 0, we reject the null hypothesis and we conclude the model is significant overall.

## Goodness of fit 

```{r}
#pearson
pearres2 = residuals(full.model,type="pearson")
pearson.tvalue = sum(pearres2^2)
c(pearson.tvalue, 1-pchisq(pearson.tvalue,7008))

#deviance
c(deviance(full.model), 1-pchisq(deviance(full.model),7008))
```



# Step model
```{r}
#both directions step model
step.model <- stepAIC(full.model, trace=0)
step.model$anova
```


# Lasso model
```{r}
#converting data to dataframe and scaling

data.matrix <- as.matrix(dat.reduced_2)
x <- model.matrix( ~ ., dat.reduced_2)

predictors <- x[,1:length(dat.reduced_2)-1]
response <-  x[,length(dat.reduced_2)]


#Using cross validation for the Lasso regression
model_lasso <- cv.glmnet(predictors, response, alpha = 1,  family = "binomial")

#Finding optimal value of lambda that minimizes cross-validation errors
plot(model_lasso)

coef(model_lasso, model_lasso$lambda.1se)
```
# Elastic model

```{r}
elastic_result <- cv.glmnet(predictors,
                        response,
                        alpha = 0.8,
                        nfolds=5,
                        type.measure="mse",
                        family="binomial",
                        standardize=FALSE)

coef(elastic_result, s = elastic_result$lambda.min)

summary(elastic_result)
```

