---
title: "Aggregated model"
author: "Marta Bras"
date: "11/25/2019"
output: html_document
---
```{r setup}
  library(knitr)
  opts_knit$set(root.dir=normalizePath('../'))
  opts_chunk$set(fig.path = "../figures/", dev='pdf') # corrected path and added dev
```

```{r, include=FALSE}
source("./Logistic Regression/DataTransformation.R")
```


# Aggregating data  - Aggregating by all categorical variables with tenure as categorical with bins defined on data transformation


```{r}
dat.reduced_3$`Churn Value` = as.numeric(dat.reduced_3$`Churn Value`)


obdata.agg.n = aggregate(`Churn Value`~ . , data = dat.reduced_3, FUN=length)


obdata.agg.y = aggregate(`Churn Value`~ . , data = dat.reduced_3, FUN=sum)


data <- cbind(obdata.agg.n, obdata.agg.y$`Churn Value`)

data <- data %>% rename_at(19,~"Total")

## Fitting the model

model.agg = glm(cbind(`Churn Value`,Total-`Churn Value`)~ .,
                data = data,family=binomial)

## summary the model
summary(model.agg)
```



## Goodness of fit 1

```{r}
## Test for overall regression
gstat = model.agg$null.deviance - deviance(model.agg)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg))-1))
```

The mdoel is not a good fit.


```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg), 1-pchisq(deviance(model.agg),40))

## Residual Analysis
res = resid(model.agg,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~`Senior Citizen`,xlab="Senior citizen",ylab = "Std residuals",data = data)
boxplot(res~Partner,xlab="Partner",ylab = "Std residuals",data = data)
boxplot(res~Dependents,xlab="Dependents",ylab = "Std residuals",data = data)
boxplot(res~`Phone Service`,xlab="Phone service",ylab = "Std residuals",data = data)
boxplot(res~`Multiple Lines`,xlab="Multiple lines",ylab = "Std residuals",data = data)
boxplot(res~`Internet Service`,xlab="Internet service",ylab = "Std residuals",data = data)
boxplot(res~Tenure,xlab="Tenure",ylab = "Std residuals",data = data)


qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```

It is not a good fit.  Dependents, internet service and tenure have an unequal distribution of residuals.

```{r}
#variable importance
library(caret)
library(pscl)
caret::varImp(model.agg) 
pscl::pR2(model.agg)["McFadden"]
```


## Improving the fit - Removing outliers

```{r}
cooksd <- cooks.distance(model.agg)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance

abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```


```{r}
#4/n = 4/5186
length(data$Gender)

a = which(cook > 0.0007713074)

newdata = data[-a,]
```



# Aggregating data  - Aggregating by all categorical variables with tenure as categorical with bins defined on data transformation and without outliers as higher than 4/n

```{r}
model.agg_2 = glm(cbind(`Churn Value`,Total-`Churn Value`)~ .,
                data = newdata,family=binomial)
  summary(model.agg_2)

```




## Goodness of fit 2

```{r}
## Test for overall regression
gstat = model.agg_2$null.deviance - deviance(model.agg_2)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg_2))-1))
```

Not a good fi

```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg_2,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg_2), 1-pchisq(deviance(model.agg_2),40))

## Residual Analysis
res = resid(model.agg_2,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Gender",ylab = "Std residuals",data = newdata)
boxplot(res~`Senior Citizen`,xlab="Senior citizen",ylab = "Std residuals",data = newdata)
boxplot(res~Partner,xlab="Partner",ylab = "Std residuals",data = newdata)
boxplot(res~Dependents,xlab="Dependents",ylab = "Std residuals",data = newdata)
boxplot(res~`Phone Service`,xlab="Phone service",ylab = "Std residuals",data = newdata)
boxplot(res~`Multiple Lines`,xlab="Multiple lines",ylab = "Std residuals",data = newdata)
boxplot(res~`Internet Service`,xlab="Internet service",ylab = "Std residuals",data = newdata)
boxplot(res~Tenure,xlab="Tenure",ylab = "Std residuals",data = newdata)

qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```

Not normal and from the histograms we can see not a good fit. 

# Aggregating data  - Aggregating by all categorical variables with tenure as categorical with bins defined on data transformation and without outliers as higher than 4/n and removing tenure, dependents and multiple lines

```{r}
#removing tenure
drop <- c("Tenure", "`Internet Service`", "`Multiple Lines`")
newdata_2 = newdata[,!(names(newdata) %in% drop)]

```

```{r}
model.agg_3 = glm(cbind(`Churn Value`,Total-`Churn Value`)~ .,
                data = newdata_2,family=binomial)
summary(model.agg_3)

```




## Goodness of fit 2

```{r}
## Test for overall regression
gstat = model.agg_3$null.deviance - deviance(model.agg_3)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg_3))-1))
```


```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg_3,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg_2), 1-pchisq(deviance(model.agg_2),40))

## Residual Analysis
res = resid(model.agg_3,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Age Group",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Senior Citizen`,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~Partner,xlab="Gender",ylab = "Std residuals",data = newdata_2)
#boxplot(res~Dependents,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Phone Service`,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Multiple Lines`,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Internet Service`,xlab="Gender",ylab = "Std residuals",data = newdata_2)



qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```

## checking outliers again
```{r}
cooksd <- cooks.distance(model.agg_3)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance

abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```

## removing outliers
```{r}
#4/n = 4/4870
length(newdata_2$Gender)

a = which(cook > 0.0008213552)
newdata_3 = newdata_2[-a,]
```


## running again with no outleirs
```{r}
model.agg_4 = glm(cbind(`Churn Value`,Total-`Churn Value`)~ .,
                data = newdata_3,family=binomial)
summary(model.agg_4)

```




## Goodness of fit 2

```{r}
## Test for overall regression
gstat = model.agg_4$null.deviance - deviance(model.agg_4)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg_4))-1))
```


```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg_4,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg_4), 1-pchisq(deviance(model.agg_4),40))

## Residual Analysis
res = resid(model.agg_4,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Age Group",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Senior Citizen`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~Partner,xlab="Gender",ylab = "Std residuals",data = newdata_3)
#boxplot(res~Dependents,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Phone Service`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Multiple Lines`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Internet Service`,xlab="Gender",ylab = "Std residuals",data = newdata_3)



qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```

Still not a good fit





# Step model 
```{r}
#both directions step model
step.model <- stepAIC(model.agg_4, trace=0)
step.model$anova
```


```{r}
model.agg_5 = glm(cbind(`Churn Value`,Total-`Churn Value`)~Partner + `Internet Service` + 
    `Online Security` + `Online Backup` + `Tech Support` + `Streaming TV` + 
    `Streaming Movies` + Contract + `Paperless Billing` + `Payment Method`, data = newdata_3,family=binomial)
summary(model.agg_5)

```


## Goodness of fit 2

```{r}
## Test for overall regression
gstat = model.agg_5$null.deviance - deviance(model.agg_5)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg_5))-1))
```


```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg_5,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg_4), 1-pchisq(deviance(model.agg_4),40))

## Residual Analysis
res = resid(model.agg_4,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Age Group",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Senior Citizen`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~Partner,xlab="Gender",ylab = "Std residuals",data = newdata_3)
#boxplot(res~Dependents,xlab="Gender",ylab = "Std residuals",data = newdata_2)
boxplot(res~`Phone Service`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Multiple Lines`,xlab="Gender",ylab = "Std residuals",data = newdata_3)
boxplot(res~`Internet Service`,xlab="Gender",ylab = "Std residuals",data = newdata_3)



qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```







# Aggregating data  - Aggregating by all categorical variables with tenure as categorical with bins defined on data transformation

```{r}
dat.reduced_3$`Churn Value` = as.numeric(dat.reduced_3$`Churn Value`)


obdata.agg.n = aggregate(`Churn Value`~ . , data = dat.reduced_3, FUN=length)


obdata.agg.y = aggregate(`Churn Value`~ . , data = dat.reduced_3, FUN=sum)

Total <- obdata.agg.y$`Churn Value`

data <- cbind(obdata.agg.n, Total)

## Fitting the model

model.agg = glm(cbind(`Churn Value`,difference)~ .,
                data = data,family=binomial)

## summary the model
summary(model.agg)
```



## Goodness of fit 1

```{r}
## Test for overall regression
gstat = model.agg$null.deviance - deviance(model.agg)
cbind(gstat, 1-pchisq(gstat,length(coef(model.agg))-1))
```

The mdoel is not a good fit


```{r}
## Test for GOF: Using deviance residuals
deviances2 = residuals(model.agg,type="deviance")
dev.tvalue = sum(deviances2^2)
c(dev.tvalue, 1-pchisq(dev.tvalue,40))
#OR
c(deviance(model.agg), 1-pchisq(deviance(model.agg),40))

## Residual Analysis
res = resid(model.agg,type="deviance")
par(mfrow=c(2,2))
boxplot(res~Gender,xlab="Age Group",ylab = "Std residuals",data = data)
boxplot(res~`Senior Citizen`,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~Partner,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~Dependents,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~`Phone Service`,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~`Multiple Lines`,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~`Internet Service`,xlab="Gender",ylab = "Std residuals",data = data)
boxplot(res~Tenure,xlab="Gender",ylab = "Std residuals",data = data)


qqnorm(res, ylab="Std residuals")
qqline(res,col="blue",lwd=2)
hist(res,10,xlab="Std residuals", main="")
```

